Theory about Normal Basis on $F(2^k)$

### 2017-03-03 dissertation/submission/recha.tex
after submission: revision 2

\chapter{Word-Level Traversal of Finite State Machines using Algebraic Geometry}
\label{ch:reacha}
Reachability analysis is a basic component of sequential circuit verification, especially for 
formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
tools, in order to improve various performance indicators such as latency, clock skew or power
density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating} are applied.
% some major refactoring and attachments are made on original designs. 
These modifications may introduce bugs, errors or malfunctions to the original logic and cause problems. 
Based on traditional localized simulation or formal verification method ({\it e.g.} equivalence checking), 
designers are reluctant to make aggressive optimization
since the malfunctions are considered as ``faults" in the circuits.
However, if the circuit behavior is carefully investigated, it may become evident that 
those ``faults" may never be activated during a restricted execution starting 
from legal initial states and with legal inputs. Thus we will call those ``faults" as 
{\it spurious faults} (false negatives), since they will not affect the circuit's normal behavior.

Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
If we apply constraints upon the machine to make it start from designated initial states, and 
take specific legal inputs, a set of reachable states can be derived. 
As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
spurious faults by checking if they belong to the unreachable states. From the spurious fault validation 
perspective, reachability analysis is a must when developing full set of sequential circuit verification
techniques.

There are quite a few methods to perform reachability checking on FSMs. One among them is 
state space traversal. Conventionally, the algorithm is based on bit-level techniques such as
binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
which brings critical advantages. In this chapter the approach will be described and discussed in depth, 
with examples and experiments showing its feasibility when applied on general circuit benchmarks.

\section{Motivation}
Before introducing the details of our approach, there are a few questions to ask: what is 
the benefit of executing FSM traversal at word-level? How can algebraic geometry make this happen?
The answers can be found in this section, as a statement of the motivation of our research.

\subsection{FSM Traversal Algorithms}
Sequential circuits are modeled as FSMs, which can be implemented as graphs. Thus a 
graph-traversal based algorithm is created to analyze the reachable states \cite{coudert2003unified}.
A traversal algorithm using the concept of implicit state enumeration is proposed \cite{cho1993redundancy}.
Concretely, the algorithm is given in Algorithm \ref{alg:BFS}:

@@

Above algorithm describes a breath-first-search (BFS) traversal in
state space. The traversal algorithm is a simple variation of BFS algorithm where 
states are nodes and transitions are arcs. Each state is uniquely encoded by a combination
of a set of register data, which is usually represented by a Boolean vector.

Since a typical sequential circuit usually contains a combinational logic
component, the traversal algorithm analyzes the combinational logic and derives the transition 
function for one-step reachability within current time-frame, and extends the result to complete 
execution paths through unrolling. If each state encoding ({\it i.e.} exact values in the selected registers) is
explicitly analyzed and counted during the unrolling 
procedure, this unrolling is called {\bf explicit unrolling}. In the BFS traversal algorithm, the states cannot be 
directly read in the execution; instead, they are implicitly represented using a conjunction of several 
Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.

However, the BFS traversal algorithm proposed by the author is usually not practical. The conjunctions of 
Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
the size of BDD explodes when the formulas become too long and too complicated. In \cite{cho1993redundancy},
the authors make a compromise between accuracy and cost, and turn to approximate reachability. 
In this research work, the aim is to explore a word-level technique which can make a accurate reachability
analysis available.

\subsection{Word-Level Data Flow on Modern Datapath Designs}
The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
implementation of large datapaths possible. In recent decades, 64-bit or even larger datapaths frequently appear in 
modern digital IC designs such as powerful central processing units (CPUs) and high bandwidth memory (HBM).
Meanwhile, with the development of electronic design automation (EDA) tools, data flow 
is described by the designer as word-level specifications. Therefore, it will be straightforward 
and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
techniques will greatly reduce the state space and make verification more efficient.

In order to throw light on the advantages using word-level techniques, we pick a typical 
digital circuit component in modern 64-bit MIPS processor as an example.

\begin{Example}
Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
MIPS. Initially, one multiplicand is preloaded to the lower 64 bits of the product registers. 
Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
to add on the other multiplicand. For each iteration the data in product registers shifts right by 1 bit.
Finally when the most significant bit (MSB) of preloaded multiplicand arrives at the MSB of product registers,
the registers contains the result -- 128 bits product. The behavior can be described by the following algorithm:

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
\caption{A 64-bit sequential multiplication hardware design.}
\label{fig:MIPS}}
\end{figure}

% List the detailed bit-level property need to check P==P+B
Traditionally, to verify the functional correctness of this multiplier, satisfiability (SAT) based or BDD based 
model checking is applied on basic function units. For example, as a part of functional verification, 
we would like to check ``$P=P_{1/2}+B$" is correctly executed. Then in a model checker we need to add following 
specifications:

@@

We can see that when checking a single part of the whole structure, the number of clauses 
needed will increase to $k+1$ when using $k$-bit datapath. Considering the formula representing 
carry-in will become longer and longer, the final conjunction of all clauses will contain 
$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables:
\begin{equation}
\label{eqn:wordspec}
S=P_{1/2}+B
\end{equation}

The abstraction to word-level will reduce symbolic storage and execution cost; the complexity 
to traverse the state space will be greatly reduced.
\end{Example}

On the other hand, when implementation details of the datapath are not available, it is not convenient 
any more for users of conventional model checker. The reason is that the user has to write all clauses 
for the implementation, which contains cross-literals, {\it e.g.} $s_2$ may associate with $a_1, p_0$, {\it etc.} If the user 
is not familiar with the implementation of this adder, those cross-literals will bring
confusions. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
the verification tool will be very user-friendly and straightforward even if the implementation details
are in a black box.

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=4in]{newfig/Lagrange.pdf}
\caption{Gate-level netlist for Lagrange's interpolation example.}
\label{fig:Lagrange}}
\end{figure}

For each element in $\F_{2^3}$, we write down the truth table as Table \ref{tab:truthtable}.
Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
$Z = \Func(A)$. Recall Lagrange's interpolation formula:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}

\begin{table}[bp]
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\end{table}

@@

Based on the aforementioned discussions, the importance of performing FSM traversal at word-level 
can bring a dimension of abstraction in sequential circuit reachability analysis.
To overcome the cost incurred by searching in a large space, we propose to use word-level 
polynomials to represent the states and transition relations. As a result, states are categorized into 
sets represented by a small number of word-level polynomials (more specifically, the varieties to the polynomial ideals),
and multiple transition relations are therefore merged together. All of these efforts reduce the cost of 
state space, meanwhile lowering the time complexity to traverse such a state space. As Lagrange's interpolation
confirms the existence of word-level abstraction for bit-level circuits, the concept
can be extended to encode the state space of sequential circuit by word-level polynomials in $\Fkk$.
% we prove the feasibility of word-level techniques on general circuits' verification.
% Thus we cover both the necessity and sufficiency of developing such 
% a word-level traversal technique in our work.
% They are also answers to the questions at the beginning of this section.

\section{FSM Reachability using Algebraic Geometry}
\label{sec:reach}
% \subsection{An example illustrating our proposed approach}
We use symbolic state reachability with algebraic
geometry concepts. It is an abstraction based on word operand
definition of datapaths in circuits, and it can be applied
to arbitrary FSMs by bundling a set of bit-level variables together as
one or several word-level variables.  The abstraction polynomial,
encoding the reachable state space of the FSM, is obtained through
computing a GB over $\Fkk$ of the polynomials of the circuit using an
elimination term order based on Theorem \ref{thm:elimth}.

\subsection{FSM Model for Sequential Circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, {\it i.e.} the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i$: 
$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$

@@

The {\it transition function} $\Delta$ is given by Boolean equations
  of the flip-flops of the circuit: $t_i = \Delta_i(s, x)$, where
  $t_i$ is a next state variable, $s$ represents the present   state
  variables and $x$ represents the input variables. The {\it
    transition relation of the FSM} is then represented as:  
\begin{equation} 
T(s, x, t) =   \prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)
\end{equation}
where $n$ is the number of flip flops, and $\overline{\oplus}$ is XNOR
operation. Let $from$ denote the set of initial states, then the
image of the initial states, under the transition function $\Delta$ is
finally computed as:
\begin{align}
\label{eqn:img}
to = \text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
  \cdot from ] 
%= \exists _s ~\exists _x ~\prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)\cdot from
\end{align}

@@

\begin{Example}
For the circuit in Figure \ref{fig:fsm}(b), we have the transition
functions of the machine as:

\begin{align*}
\Delta_1: & ~t_0 \overline{\oplus} ((\overline{x \vee s_0 \vee s_1}) \vee s_0 s_1)\\
\Delta_2: & ~t_0 \overline{\oplus} (\overline{s_0}x \vee \overline{s_1}s_0)\\
from:     & ~from^0 = s_0\cdot s_1
\end{align*}

@@

Our objective is to model the transition functions $\Delta$ as a
polynomial ideal $J$, and to perform the image computations 
%(Algorithm \ref{alg:BFS}, line 4) 
using Gr\"obner bases over Galois fields. {\it
This requires to perform quantifier elimination; which can be
accomplished using the GB computation over $\Fkk$ using elimination
ideals} \cite{gao:qe-gf-gb}. Finally, the set union, intersection and
complement operations are also to be implemented in algebraic geometry.

@@

{\bf Set operations:} In Lines 5 and 6 of
Algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
\textbf{intersection} and \textbf{complement} of varieties over
$\mathbb{F}_{2^k}$, for which we again use algebraic geometry concepts.

@@

Now we make an assumption that $x \in V(g)$. Then $g(x) = 0 =
d(x)g_1(x)$ which means either $h(x) = 0$ or $g_1 (x) = 0$.

@@

Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$
{\it i.e.} the variety of vanishing ideal contains
 all possible valuations of variables, so it constitutes the {\bf
   universal set}. Subsequently, based on Theorem \ref{thm:quotient},
 the {\bf absolute complement} $V(J')$ of a variety $V(J)$ can be
 computed as:

@@

\section{Problem Setup and Formulation}
\begin{Problem}
We use the notions from Figure \ref{fig:seqmodel}(a) in this setup. 
\begin{enumerate}[{1)}] 
\item The circuit is modeled over $\Fkk = \F_2[x]\pmod{p(x)}, p(\alpha) = 0$, where $k$ is the number of flip-flops,
	the PS variables $\{s_0,\dots,s_{k-1}\}$, NS variables $\{t_0,\dots,t_{k-1}\}$.
\item Denote $S$ as the PS word-level by the following polynomial:
$$f_S: S = s_0+s_1\alpha+s_2\alpha^2+\cdots+s_{k-1}\alpha^{k-1}$$
Similarly we define NS word-level variable $T$ by:
$$f_T: T = t_0+t_1\alpha+t_2\alpha^2+\cdots+t_{k-1}\alpha^{k-1}$$
\item Impose ATO for sequential circuits which is 
$$LEX: \text{ all bit-level variables in any order }>S>T$$
\item Write polynomials $f_1,\dots,f_s$ for each gate, and construct ideal describing the circuit:
$$J_{ckt} = \langle f_1,\dots,f_s,f_S,f_T\rangle$$
as well as the ideal with vanishing polynomials:
$$J_0 = \langle \dots,x_i^2-x_i,\dots,S^{2^k}-S,T^{2^k}-T\rangle$$
where $x_i$ corresponds to bit-level variables denoting wires in the circuit.
\item Compute GB with ATO for $G$. Then obtain the projection of variety only on NS variable $T$, by
$$G\cap \Fkk[T]$$
\end{enumerate}
\end{Problem}

@@

Since the formerly reached state ``$reached = T$'', its complement is
computed using Corollary \ref{cor:complement} 
$$\langle T^4-T\rangle:\langle T\rangle
= \langle T^3+1\rangle.$$

$V(\langle T^3 + 1\rangle) = \{1, \alpha, \alpha+1\}$ denoting the states
$\{01,10,11\}$. Then the newly reached state set in this iteration is 
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle
T^2+(\alpha+1)T+\alpha \rangle$$ We add these states 
to formerly reached states 
\begin{align*}
reach &= \langle T\rangle \cdot \langle T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T^3+(\alpha+1)T^2+\alpha T\rangle
\end{align*}
 {\it i.e.} states $\{00,01,10\}$. We update the present states
for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$

In the second iteration, we compute the reduced GB with the same term
order for ideal $J = J_{ckt}+J_v+\langle from^1\rangle$. 
It includes a polynomial generator $$to^2 = T^2+\alpha T$$ denotes states
$\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
= \langle T + 1+\alpha\rangle$$ ({\it i.e.} states $\{11\}$). We compute the newly reached state 
$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$

@@

The remainder from $Spoly$ reduction contains bit-level PS variables, and our objective is to get a polynomial 
containing only word-level PS variables. One possible method is 
to rewrite bit-level variables in term of word-level variables, {\it i.e.}
\begin{equation}
\label{eqn:BLVS_reacha}
s_i = \mathcal{G}(S)
\end{equation}

>>>>

\chapter{Word-Level Traversal of Finite State Machines using Algebraic Geometry}
\label{ch:reacha}
Reachability analysis is a basic component of sequential circuit verification, especially for 
formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
tools, in order to improve various performance indicators such as latency, clock skew, or power
density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
sequential power optimization \cite{mathur2009power}, and clock-gating techniques \cite{clockgating} are applied.
% some major refactoring and attachments are made on original designs. 
These modifications may introduce bugs, errors or malfunctions to the original logic and cause problems. 
Based on traditional localized simulation or formal verification method (e.g., equivalence checking), 
designers are reluctant to make aggressive optimization
since the malfunctions are considered ``faults" in the circuits.
However, if the circuit behavior is carefully investigated, it may become evident that 
those faults may never be activated during a restricted execution starting 
from legal initial states and with legal inputs. Thus we will call those faults
{\it spurious faults} (false negatives), since they will not affect the circuit's normal behavior.

Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
If we apply constraints upon the machine to make it start from designated initial states, and 
take specific legal inputs, a set of reachable states can be derived. 
As long as the faults can be modeled as bad states, we can judge whether they are 
spurious faults by checking if they belong to the unreachable states. From the spurious fault validation 
perspective, reachability analysis is a must when developing a full set of sequential circuit verification
techniques.

There are quite a few methods to perform reachability checking on FSMs. One of them is 
state space traversal. Conventionally, the algorithm is based on bit-level techniques such as
binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm at word level,
which brings critical advantages. In this chapter the approach will be described and discussed in depth, 
with examples and experiments showing its feasibility when applied on general circuit benchmarks.

\section{Motivation}
Before introducing the details of our approach, there are a few questions to ask: What is 
the benefit of executing FSM traversal at word level? How can algebraic geometry make this happen?
The answers can be found in this section, as a statement of the motivation of our research.

\subsection{FSM Traversal Algorithms}
Sequential circuits are modeled as FSMs, which can be implemented as graphs. Thus a 
graph-traversal-based algorithm is created to analyze the reachable states \cite{coudert2003unified}.
A traversal algorithm using the concept of implicit state enumeration is proposed \cite{cho1993redundancy}.
Concretely, the algorithm is given in Algorithm \ref{alg:BFS}:

@@

The above algorithm describes a breath-first-search (BFS) traversal in
state space. The traversal algorithm is a simple variation of a BFS algorithm where 
states are nodes and transitions are arcs. Each state is uniquely encoded by a combination
of a set of register data, which is usually represented by a Boolean vector.

Since a typical sequential circuit usually contains a combinational logic
component, the traversal algorithm analyzes the combinational logic and derives the transition 
function for one-step reachability within the current time-frame, and extends the result to complete 
execution paths through unrolling. If each state encoding (i.e., exact values in the selected registers) is
explicitly analyzed and counted during the unrolling 
procedure, this unrolling is called {\bf explicit unrolling}. In the BFS traversal algorithm, the states cannot be 
directly read in the execution; instead, they are implicitly represented using a conjunction of several 
Boolean formulas. Such techniques differ from explicit unrolling and are called {\bf implicit unrolling}.

However, the BFS traversal algorithm proposed by the author is usually not practical. The conjunctions of 
Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
the size of a BDD explodes when the formulas become too long and too complicated. In \cite{cho1993redundancy},
the authors make a compromise between accuracy and cost, and turn to approximate reachability. 
In this research work, the aim is to explore a word-level technique which can make an accurate reachability
analysis available.

\subsection{Word-Level Data Flow on Modern Datapath Designs}
The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
integrates 3.3 billion transistors on a $125~\text{mm}^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
implementation of large datapaths possible. In recent decades, 64-bit or even larger datapaths frequently appear in 
modern digital IC designs such as powerful central processing units (CPUs) and high-bandwidth memory (HBM).
Meanwhile, with the development of electronic design automation (EDA) tools, data flow 
is described by the designer as word-level specifications. Therefore, it will be straightforward 
and beneficial for users if formal verification tools can work at word level. Moreover, adopting word-level 
techniques will greatly reduce the state space and make verification more efficient.

In order to throw light on the advantages of using word-level techniques, we pick a typical 
digital circuit component in a modern 64-bit MIPS processor as an example.

\begin{Example}
Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
MIPS. Initially, one multiplicand is preloaded to the lower 64 bits of the product registers. 
Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
to add on the other multiplicand. For each iteration the data in product registers shifts right by 1 bit.
Finally when the most significant bit (MSB) of preloaded multiplicand arrives at the MSB of product registers,
the registers contains the result -- a 128-bit product. The behavior can be described by the following algorithm:

@@

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
\caption{A 64-bit sequential multiplication hardware design.}
\label{fig:MIPS}}
\end{figure}

% List the detailed bit-level property need to check P==P+B
Traditionally, to verify the functional correctness of this multiplier, SAT-based or BDD-based 
model checking is applied on basic function units. For example, as a part of functional verification, 
we would like to check ``$P=P_{1/2}+B$" is correctly executed. Then in a model checker we need to add following 
specifications:

@@

We can see that when checking a single part of the whole structure, the number of clauses 
needed will increase to $k+1$ when using $k$-bit datapath. Considering the formula representing 
carry-in will become longer and longer, the final conjunction of all clauses will contain 
$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables:
\begin{equation}
\label{eqn:wordspec}
S=P_{1/2}+B
\end{equation}
the abstraction to word-level will reduce symbolic storage and execution cost; the complexity 
to traverse the state space will be greatly reduced.
\end{Example}

On the other hand, when implementation details of the datapath are not available, it is not convenient 
any more for users of conventional model checkers. The reason is that the user has to write all clauses 
for the implementation, which contains cross-literals, e.g., $s_2$ may associate with $a_1, p_0$, etc. If the user 
is not familiar with the implementation of this adder, those cross-literals will bring
confusion. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
the verification tool will be very user-friendly and straightforward even if the implementation details
are in a black box.

@@

\begin{figure}[tbp]
\centering{
\includegraphics[width=4in]{newfig/Lagrange.pdf}
\caption{Gate-level netlist for Lagrange's interpolation example.}
\label{fig:Lagrange}}
\end{figure}

For each element in $\F_{2^3}$, we write down the truth table as Table \ref{tab:truthtable}.
Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, i.e., 
$Z = \Func(A)$. Recall Lagrange's interpolation formula:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}

\begin{table}[tbp]
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\end{table}

@@

Based on the aforementioned discussions, the importance of performing FSM traversal at word level 
can bring a dimension of abstraction in sequential circuit reachability analysis.
To overcome the cost incurred by searching in a large space, we propose to use word-level 
polynomials to represent the states and transition relations. As a result, states are categorized into 
sets represented by a small number of word-level polynomials (more specifically, the varieties to the polynomial ideals),
and multiple transition relations are therefore merged together. All of these efforts reduce the cost of 
state space, meanwhile lowering the time complexity to traverse such a state space. As Lagrange's interpolation
confirms the existence of word-level abstraction for bit-level circuits, the concept
can be extended to encode the state space of a sequential circuit by word-level polynomials in $\Fkk$.
% we prove the feasibility of word-level techniques on general circuits' verification.
% Thus we cover both the necessity and sufficiency of developing such 
% a word-level traversal technique in our work.
% They are also answers to the questions at the beginning of this section.

\section{FSM Reachability Using Algebraic Geometry}
\label{sec:reach}
% \subsection{An example illustrating our proposed approach}
We use symbolic state reachability with algebraic
geometry concepts. It is an abstraction based on word operand
definition of datapaths in circuits, and it can be applied
to arbitrary FSMs by bundling a set of bit-level variables together as
one or several word-level variables.  The abstraction polynomial,
encoding the reachable state space of the FSM, is obtained through
computing a GB over $\Fkk$ of the polynomials of the circuit using an
elimination term order based on Theorem \ref{thm:elimth}.

\subsection{FSM Model for Sequential Circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is called a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, i.e., the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i$: 
$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$

@@

The {\it transition function} $\Delta$ is given by Boolean equations
  of the flip-flops of the circuit: $t_i = \Delta_i(s, x)$, where
  $t_i$ is a next state variable, $s$ represents the present   state
  variables and $x$ represents the input variables. The {\it
    transition relation of the FSM} is then represented as  
\begin{equation} 
T(s, x, t) =   \prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)
\end{equation}
where $n$ is the number of flip flops, and $\overline{\oplus}$ is XNOR
operation. Let $from$ denote the set of initial states, then the
image of the initial states, under the transition function $\Delta$ is
finally computed as
\begin{align}
\label{eqn:img}
to = \text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
  \cdot from ] 
%= \exists _s ~\exists _x ~\prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)\cdot from
\end{align}

@@

\begin{Example}
For the circuit in Figure \ref{fig:fsm}(b), we have the transition
functions of the machine as
\begin{align*}
\Delta_1: & ~t_0 \overline{\oplus} ((\overline{x \vee s_0 \vee s_1}) \vee s_0 s_1)\\
\Delta_2: & ~t_0 \overline{\oplus} (\overline{s_0}x \vee \overline{s_1}s_0)\\
from:     & ~from^0 = s_0\cdot s_1
\end{align*}

@@

Our objective is to model the transition functions $\Delta$ as a
polynomial ideal $J$, and to perform the image computations 
%(Algorithm \ref{alg:BFS}, line 4) 
using Gr\"obner bases over Galois fields. {\it
This requires performing quantifier elimination, which can be
accomplished using the GB computation over $\Fkk$ using elimination
ideals} \cite{gao:qe-gf-gb}. Finally, the set union, intersection, and
complement operations are also to be implemented in algebraic geometry.

@@

{\bf Set operations:} In Lines 5 and 6 of
Algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
\textbf{intersection}, and \textbf{complement} of varieties over
$\mathbb{F}_{2^k}$, for which we again use algebraic geometry concepts.

@@

Now we make an assumption that $x \in V(g)$. Then $g(x) = 0 =
d(x)g_1(x)$, which means either $h(x) = 0$ or $g_1 (x) = 0$.

@@

Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$,
i.e., the variety of vanishing ideal contains
 all possible valuations of variables, so it constitutes the {\bf
   universal set}. Subsequently, based on Theorem \ref{thm:quotient},
 the {\bf absolute complement} $V(J')$ of a variety $V(J)$ can be
 computed as:

@@

\section{Problem Setup and Formulation}
\begin{Problem}
We use the notions from Figure \ref{fig:seqmodel}(a) in this setup. 
\begin{enumerate}[{1)}] 
\item The circuit is modeled over $\Fkk = \F_2[x]\pmod{p(x)}, p(\alpha) = 0$, where $k$ is the number of flip-flops,
	the PS variables $\{s_0,\dots,s_{k-1}\}$, NS variables $\{t_0,\dots,t_{k-1}\}$.
\item Denote $S$ as the PS word-level by the following polynomial:
$$f_S: S = s_0+s_1\alpha+s_2\alpha^2+\cdots+s_{k-1}\alpha^{k-1}$$
Similarly we define NS word-level variable $T$ by:
$$f_T: T = t_0+t_1\alpha+t_2\alpha^2+\cdots+t_{k-1}\alpha^{k-1}$$
\item Impose ATO for sequential circuits, which is 
$$LEX: \text{ all bit-level variables in any order }>S>T$$
\item Write polynomials $f_1,\dots,f_s$ for each gate, and construct ideal describing the circuit:
$$J_{ckt} = \langle f_1,\dots,f_s,f_S,f_T\rangle$$
as well as the ideal with vanishing polynomials:
$$J_0 = \langle \dots,x_i^2-x_i,\dots,S^{2^k}-S,T^{2^k}-T\rangle$$
where $x_i$ corresponds to bit-level variables denoting wires in the circuit.
\item Compute GB with ATO for $G$. Then obtain the projection of variety only on NS variable $T$, by
$$G\cap \Fkk[T]$$
\end{enumerate}
\end{Problem}

@@

Since the formerly reached state ``$reached = T$," its complement is
computed using Corollary \ref{cor:complement} 
$$\langle T^4-T\rangle:\langle T\rangle
= \langle T^3+1\rangle.$$

$V(\langle T^3 + 1\rangle) = \{1, \alpha, \alpha+1\}$ denoting the states
$\{01,10,11\}$. Then the newly reached state set in this iteration is 
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle
T^2+(\alpha+1)T+\alpha \rangle$$ We add these states 
to formerly reached states 
\begin{align*}
reach &= \langle T\rangle \cdot \langle T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T^3+(\alpha+1)T^2+\alpha T\rangle
\end{align*}
i.e., states $\{00,01,10\}$. We update the present states
for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$

In the second iteration, we compute the reduced GB with the same term
order for ideal $J = J_{ckt}+J_v+\langle from^1\rangle$. 
It includes a polynomial generator $$to^2 = T^2+\alpha T$$ denotes states
$\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
= \langle T + 1+\alpha\rangle$$ (i.e., states $\{11\}$). We compute the newly reached state 
$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$

@@

The remainder from $Spoly$ reduction contains bit-level PS variables, and our objective is to get a polynomial 
containing only word-level PS variables. One possible method is 
to rewrite bit-level variables in terms of word-level variables, i.e.,
\begin{equation}
\label{eqn:BLVS_reacha}
s_i = \mathcal{G}(S)
\end{equation}
